---
title: "STOR 320 Tutorial 11"
author: "Mario Giacomazzo"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=F)
options(scipen=999)
library(tidyverse)    #Essential Functions
library(modelr)       #Helpful Functions in Modeling
DATA=read_csv("AirWaterTemp.csv",col_types=cols()) #River Data
```

# Introduction

Today, we will work with daily water temperature and air temperature data observed for `r length(unique(DATA$L))` rivers in Spain. The goal of this tutorial is to identify the best model for predicting the maximum water temperature given the maximum air temperature. In the preview below, `W` represents the daily maximum water temperature and `A` represents the daily maximum air temperature. The data contains almost a full year of data for each of the `r length(unique(DATA$L))` different rivers.

```{r,echo=F}
head(DATA)
```


#Part 1: Examining the Relationship

##Chunk 1: Overall Relationship
```{r,echo=F,eval=T,message=F}
ggplot(data=DATA) +
  geom_point(aes(x=A,y=W),alpha=0.3)+
  geom_smooth(aes(x=A,y=W)) +
  theme_minimal()
```


##Chunk 2: Location-Specific Relationship

```{r,echo=F,eval=F}
Location = 103

DATA %>% filter(L==Location) %>%
  ggplot()+
    geom_point(aes(x=A,y=W),alpha=0.3)+
    geom_smooth(aes(x=A,y=W)) +
    theme_minimal()
  
```

##Chunk 3: Split Data into Train and Test Sets
```{r,eval=F}
set.seed(INTEGER)
TEST.LOCATIONS=sample(x=unique(DATA$L),size=3,replace=F)

TRAIN = anti_join(DATA,tibble(L=TEST.LOCATIONS),by="L")
TEST = semi_join(DATA,tibble(L=TEST.LOCATIONS),by="L")
```

##Chunk 4: Plots of Relationship for Train and Test Data
```{r,echo=F,eval=F}
WAPLOT2.func=function(DATA){
  ggplot(data=DATA)+
    geom_point(aes(x=A,y=W),alpha=0.3)+
    geom_smooth(aes(x=A,y=W)) +
    theme_minimal()
}

WAPLOT2.func(TRAIN)
WAPLOT2.func(TEST)
```


#Part 2: Linear Regression Model

##Chunk 1: Fitting Linear Model to Train Data
```{r,echo=F,eval=F}
linmod=lm(W~A,data=TRAIN)
summary(linmod)
```

##Chunk 2: Getting Predictions from Linear Model
```{r,eval=F}
TRAIN2 = TRAIN %>% add_predictions(linmod,var="linpred")
TEST2 = TEST %>% add_predictions(linmod,var="linpred")
```

##Chunk 3: Getting Residuals from Linear Model
```{r,eval=F}
TRAIN3 = TRAIN2 %>% add_residuals(linmod,var="linres")
TEST3 = TEST2 %>% add_residuals(linmod,var="linres")
```

#Part 3: Polynomial Regression Model

##Chunk 1: Fitting Polynomial Regression Models
```{r,eval=F}
poly2mod=lm(W~A+I(A^2),data=TRAIN)
poly3mod=lm(W~A+I(A^2)+I(A^3),data=TRAIN)
poly4mod=lm(W~A+I(A^2)+I(A^3)+I(A^4),data=TRAIN)
anova(linmod,poly2mod,poly3mod,poly4mod,test="Chisq")
```

##Chunk 2: Getting Predictions from Polynomial Models
```{r,eval=F}
TRAIN4 =TRAIN3 %>% 
  add_predictions(poly2mod,var="poly2pred") %>%
  add_predictions(poly3mod,var="poly3pred") %>%
  add_predictions(poly4mod,var="poly4pred")
  
TEST4 =TEST3 %>% 
  add_predictions(poly2mod,var="poly2pred") %>%
  add_predictions(poly3mod,var="poly3pred") %>%
  add_predictions(poly4mod,var="poly4pred")
```

##Chunk 3: Getting Residuals from Polynomial Models
```{r,eval=F}
TRAIN5 =TRAIN4 %>% 
  add_predictions(poly2mod,var="poly2pred") %>%
  add_predictions(poly3mod,var="poly3pred") %>%
  add_predictions(poly4mod,var="poly4pred")
  
TEST5 =TEST4 %>% 
  add_predictions(poly2mod,var="poly2pred") %>%
  add_predictions(poly3mod,var="poly3pred") %>%
  add_predictions(poly4mod,var="poly4pred")
```

#Part 4: Nonlinear Logistic Model

##Chunk 1: Logistic Model Investigation
```{r,echo=F,eval=F}
set.seed(INTEGER)
EXAMPLE=tibble(
  x=rnorm(10000,mean=0,sd=5),
  y=7+12/(1+exp(4-1*x))
)
ggplot(data=EXAMPLE) +
  geom_point(aes(x=x,y=y)) +
  theme_minimal()
```

##Chunk 2: Essential Functions for Nonlinear Logistic Model
```{r,eval=F}
logistic.model=function(COEF,DATA){
  pred=COEF[1]+COEF[2]/(1+exp(COEF[3]-COEF[4]*DATA$A))
}
MSE.logistic=function(COEF,DATA){
  error=DATA$W-logistic.model(DATA=DATA,COEF=COEF)
  sq.error=error^2
  mse=mean(sq.error,na.rm=T)
  return(mse)
}
```

##Chunk 3: Estimate Parameters for Nonlinear Logistic Model
```{r,eval=F}
logistic.mod=optim(
  par=c(min(TRAIN$W,na.rm=T),
        max(TRAIN$W,na.rm=T)-min(TRAIN$W,na.rm=T),
        median(TRAIN$A,na.rm=T),
        1),           #Smart Starting Values
  fn=MSE.logistic,    #Function to Minimize
  DATA=TRAIN          #Required Argument
)
print(logistic.mod)
```

##Chunk 4: Obtain Predictions and Residuals
```{r,eval=F}
TRAIN6=TRAIN5 %>% mutate(logpred=logistic.model(COEF=logistic.mod$par,DATA=TRAIN5),
                         logres=W-logpred)
TEST6=TEST5 %>% mutate(logpred=logistic.model(COEF=logistic.mod$par,DATA=TEST5),
                         logres=W-logpred)
```

#Intermission:

The function `save.image()` in R can be used to save all objects in the global environment. This is very helpful when you want work off your results without rerunning all previous R code. The name of the exported information should contain the file extension *.Rdata*. These files can be extremely large depending how much RAM was utilized in your R session.  The function `load()` can be used to import a previous workspace.

For more information on *.Rdata* file types, see https://fileinfo.com/extension/rdata for help.

```{r,eval=F}
save.image("Tutorial11.Rdata")
```


#Part 5: Evaluation by Visualization

#Chunk 1: Plotting Models
```{r,echo=F,eval=F}
TEST6 %>%
  select(L,A,W,linpred,poly2pred,poly3pred,poly4pred,logpred)%>%
  gather(linpred:logpred,key="Model",value="Pred",factor_key=T) %>%
    ggplot() + 
      geom_point(aes(x=A,y=W),color="gray") + 
      theme_minimal() +
      geom_line(aes(x=A,y=Pred,color=Model),size=2)
```

#Chunk 2: Predicted Versus Actual Max Water Temperature
```{r,echo=F,eval=F}
TEST6 %>%
  select(L,A,W,linpred,poly2pred,poly3pred,poly4pred,logpred)%>%
  gather(linpred:logpred,key="Model",value="Pred",factor_key=T) %>%
    ggplot() + 
    geom_point(aes(x=W,y=Pred,color=Model)) + 
    geom_abline(a=0,b=1,color="black",size=2) +
    xlab("Maximum Water Temperature") +
    ylab("Predicted Water Temperature") +
    theme_minimal()
```

#Chunk 3: Plotting Residuals Versus Time
```{r,echo=F,eval=F}
TEST6 %>%
  select(L,A,W,TIME,linres,poly2res,poly3res,poly4res,logres)%>%
  gather(linres:logres,key="Model",value="Res",factor_key=T) %>%
    ggplot() + 
      geom_line(aes(x=TIME,y=Res),color="lightskyblue2") + 
      geom_hline(yintercept=0,color="black",linetype=2,size=1.5) +
      xlab("Time") +
      ylab("Residual") +
      theme_dark() +
      facet_grid(Model~.)
```

#Chunk 4: Evaluating The Location-Specific Error
```{r,echo=F,eval=F}
TEST6 %>%
  select(L,A,W,TIME,linpred,logpred) %>%
  gather(linpred:logpred,key="Model",value="Pred",factor_key=T) %>%
    ggplot() +
      geom_point(aes(x=A,y=W),alpha=0.2) +
      geom_line(aes(x=A,y=Pred,color=Model),size=2) +
      theme_minimal()+facet_grid(L~.)

TEST6 %>%
  select(L,A,W,TIME,linres,logres) %>%
  gather(linres:logres,key="Model",value="Res",factor_key=T) %>%
  ggplot() +
  geom_point(aes(x=A,y=Res,color=Model)) +
  geom_hline(yintercept=0) +
  theme_minimal()+facet_grid(L~.)

TEST6 %>%
  select(L,A,W,TIME,linres,logres) %>%
  gather(linres:logres,key="Model",value="Res",factor_key=T) %>%
  ggplot() +
  geom_point(aes(x=TIME,y=Res,color=Model)) +
  geom_hline(yintercept=0) +
  theme_minimal()+facet_grid(L~.)
```























