---
title: "Exploratory Data Analysis"
author: "STOR 320.(01 OR 02) Group PLACE_GROUP_NUMBER_HERE"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

# Import Data Below
# read_csv()
```

#Creator: FIRSTNAME LASTNAME

###Q1: WRITE_QUESTION_HERE
```{r}
#
```

```{r}
#
```

###Q2: WRITE_QUESTION_HERE
```{r}
#
```

```{r}
#
```

#Interpreter: FIRSTNAME LASTNAME

###Q1: WRITE_QUESTION_HERE
```{r}
#
```

```{r}
#
```

###Q2: WRITE_QUESTION_HERE
```{r}
#
```

```{r}
#
```

#Orator: FIRSTNAME LASTNAME

###Q1: WRITE_QUESTION_HERE
```{r}
#
```

```{r}
#
```

###Q2: WRITE_QUESTION_HERE
```{r}
#
```

```{r}
#
```

#Deliverer: FIRSTNAME LASTNAME

###Q1: WRITE_QUESTION_HERE
```{r}
#
```

```{r}
#
```

###Q2: WRITE_QUESTION_HERE
```{r}
#
```

```{r}
#
```

# Follow-up Questions

### New Questions Based Off Initial Investigation

- Q1: WRITE_QUESTION_HERE
- Q2: WRITE_QUESTION_HERE
- Q3: WRITE_QUESTION_HERE 
- Q4: WRITE_QUESTION_HERE

### Investigation of Follow-up Questions

GIVE WHAT 2 QUESTIONS YOU ATTEMPTED TO INVESTIGATE FURTHER IN COMPLETE SENTENCES (*Example:* Our group decided to investigate Q2 and Q4 in further detail.)

```{r}

```

```{r}

```

#Summary
















