---
title: "Lab 2 Solutions"
author: "Tom Keefe"
date: "August 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  1-3 (Pgs. 49-50); 2-4 (Pg. 51); 2,4 (Pg. 54); 1-4 (Pg. 58)

**Assigned:** Friday, August 31, 2018

**Due:** Friday, September 7, 2018 by 5:00 PM

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a html file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. 

```{r, include=FALSE}
library(tidyverse)
library(nycflights13)
```

# Chapter 3 (Pgs. 49-50)

##  Exercise 1

### a) Had an arrival delay of 2 or more hours

```{r}
filter(flights, arr_delay >= 120)
```

### b) Flew to Houston (`IAH` or `HOU`)

```{r}
filter(flights, dest=="IAH" | dest=="HOU")
```

### c) Were operated by United, American, or Delta
```{r}
filter(flights, carrier %in% c("UA", "AA", "DL"))
```

### d) Departed in summer (July, August, and September)

```{r}
filter(flights, month %in% c(7,8,9))
```

### e) Arrived more than 2 hours late, but didn't leave late

```{r}
filter(flights, arr_delay >= 120, dep_delay <= 0)
```

### f) Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
filter(flights, dep_delay >= 60, sched_arr_time+dep_delay-arr_time > 30)
```

Note, `filter(flights, dep_delay >= 60, sched_arr_time+dep_delay-arr_time > 30)` should receive half credit. The reason this doesn't quite work is cases where the scheduled arrival is before midnight, and the actual arrival is after midnight. Then the subtraction doesn't work out properly.

### g) Departed between midnight and 6 am (inclusive)

You may have noticed that this data codes midnight as 2400 instead of 0. So, the most correct answer is `filter(flights, dep_time <= 600 | dep_time==2400)`. But, the following is perfectly acceptable:

```{r}
filter(flights, dep_time <= 600)
```
##  Exercise 2
The `between` helper is a way to write (x >= a) & (x <= b) in a way that is easy to remember and is easy to read.
```{r}
?between

```
We can use `between` to rewrite the queries above that used numeric comparisons ((d) and (f)):
```{r}
filter(flights, between(month, 7, 9))

filter(flights, between(dep_time, 0, 600))
```

##  Exercise 3
```{r}
filter(flights, is.na(dep_time))

# Then nrow will give us the number of rows.
nrow(filter(flights, is.na(dep_time)))
```
There are 8255 rows with missing `dep_time`. For these rows, other missing columns are `dep_delay`, `arr_time`, `arr_delay`, and `air_time`. These rows may represent canceled flights.


# Chapter 3 (Pg. 51)

##  Exercise 2
```{r}
# Most delayed flights. Note, I'm interpreting this as the greatest values of dep_delay. Other reasonable interpretations might be based on arr_delay, or a combination of both.
arrange(flights, desc(dep_delay))

# Earliest leaving flights. This could be interpreted as the flights that left earliest in the day, or the flights that left earliest in the year, or the flights that left earliest compared to their scheduled departure time. I'll interpret it as the flights that left earliest in the day. 
arrange(flights, dep_time)
```

Does something jump out to you about this?

This output should put midnight departures at the top, but the earliest departures here are at 12:01 AM. As your data science skills mature, you will start recognizing this as an *issue with how the data was coded*. The authors of this data chose to use 2400 as midnight, instead of 0. 

##  Exercise 3
I'm interpreting this to be the distance between the airports divided by the time in the air. Another natural interpretation of "fastest" is simply `air_time`, e.g. "Raleigh to Wilmington is a fast flight". 

A vital piece of being a data scientist is being precise about what you're asking the data. The issues of interpretation in the above questions indicate how natural questions may be imprecise. 
```{r}
arrange(flights, desc(distance/air_time))
```

##  Exercise 4
Longest and shortest can be interpreted as either time or distance. I'll use distance.
```{r}
arrange(flights, desc(distance)) # longest distance

arrange(flights, distance) # shortest distance
```

# Chapter 3 (Pg. 54)

##  Exercise 2

What happens if you include  the name of a variable multiple times in a `select()` call?

Let's try!
```{r}
select(flights, c(flight, origin, origin, dest, origin))
```
Once you use a variable, it is ignored. This is useful for avoiding repeats when using the `everything()` helper, for example.

##  Exercise 4
Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change this default?

Let's try!
```{r}
select(flights, contains("TIME"))
```

By default, the select helpers ignore case. This can be overridden with `ignore.case=FALSE`:

```{r}
select(flights, contains("TIME", ignore.case=FALSE))
```

Observe that no columns were matched.

# Chapter 3 (Pg. 58)

##  Exercise 1
Elements of dep_time have the form HHMM, so 3:30 PM is 1530. We want this to be the number of minutes past midnight. First we extract the hour (e.g. 15) by "floor dividing" by 100. 1530 divided by 100 is 15.3, and the floor of 15.3 is 15. 

Then we extract the minutes past the hour by "modding" 1530 by 100. This means we ask for the remainder of 1530 when divided by 100. This remainder is 30, as desired.

Then we multiply the hours by 60, and add the minutes past the hour. 
```{r}
transmute(flights, 
          dep_hour = dep_time %/% 100, 
          dep_minute = dep_time %% 100, 
          dep_minutes_past_midnight = 60*dep_hour + dep_minute
)
transmute(flights, 
          sched_dep_hour = sched_dep_time %/% 100, 
          sched_dep_minute = sched_dep_time %% 100, 
          sched_dep_minutes_past_midnight = 60*sched_dep_hour + sched_dep_minute
)
```


##  Exercise 2

The `air_time` should be equivalent to the `arr_time - dep_time`, but I'm sure there will be problems related to how times are represented in this dataset. Let's see!

```{r}
transmute(flights, air_time, arr_time - dep_time)
```

It seems that arr_time - dep_time is greater than air_time! Consider a flight that leaves at 4:45 AM and arrives at 5:15 AM. The `air_time` should be 30 minutes. However, the `arr_time - dep_time` will be `515-445`, which is 70! Converting arr_time and dep_time to minutes past midnight as in the previous exercise should help:

```{r}
transmute(flights, 
          dep_hour = dep_time %/% 100, 
          dep_minute = dep_time %% 100, 
          dep_minutes_past_midnight = 60*dep_hour + dep_minute,
          arr_hour = arr_time %/% 100, 
          arr_minute = arr_time %% 100, 
          arr_minutes_past_midnight = 60*arr_hour + arr_minute,
          air_time,
          arr_minutes_past_midnight-dep_minutes_past_midnight
)
```


These still don't line up. Maybe there's a timezone issue? Let's look at the Newark-San Francisco flights:

```{r}
select(
  filter(flights, origin=='EWR', dest=='SFO'),
  dep_time, arr_time, air_time
)
```

As we can see, the `air_time` for EWR-SFO flights is about six hours, but `arr_time - dep_time` is about three hours. That's because there's an additional 3 hours time difference between Newark and SFO. 

To fix this entirely, we need to know how many hours to add to the arr_time by looking at the timezone of the destination airport. We would need to "join" this data to our `flights` table, which we haven't yet covered.


##  Exercise 3
```{r}
select(flights, dep_time, sched_dep_time, dep_delay)
```

The dep_time should equal `sched_dep_time + dep_delay` (keeping in mind that flights that leave early have negative `dep_delay`).

##  Exercise 4
Note, we already did this the easy way in an earlier exercise when we used `arrange` to sort by dep_delay. Now we have to use a ranking function explicitly. 
```{r}
flights_with_delays <- mutate(flights, top_delay=dense_rank(-dep_delay))
filter(flights_with_delays, top_delay <= 10)
```

Let's just make sure that this lines up with what `arrange` produces:

```{r}
arrange(flights, desc(dep_delay))
```

As desired, these are the same flights. 