---
title: "Lab 4"
author: "FIRSTNAME LASTNAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  2,3 (Pg. 151); 2,4 (Pg. 156); 1,2 (Pgs. 160-161); 2 (Pg. 163); 2,3,4 (Pg. 168), Open Response

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
```

# Chapter 9 (Pg. 151)

##  Exercise 2
```{r}
#
```

##  Exercise 3
```{r}
#
```

# Chapter 9 (Pg. 156)

##  Exercise 2
```{r}
#
```

##  Exercise 4
```{r}
#
```

# Chapter 9 (Pgs. 160-161)

##  Exercise 1
```{r}
#
```

##  Exercise 2
```{r}
#
```

# Chapter 9 (Pg. 163)

##  Exercise 2
```{r}
#
```

# Chapter 9 (Pg. 168)

##  Exercise 2
```{r}
#
```

##  Exercise 3
```{r}
#
```

##  Exercise 4
```{r}
#
```


# Open Response

For this exercise, I want you to use the dataset your group selected for your final project and create 4 plots (bar plot, box plot, histogram, scatter plot). You should use the `ggplot()` function with different geoms for all figures. All the plots should use different variables (**4 total variables**), and these plots can be used in your exploratory data analysis.

## Step 1: Bar plot
```{r}
#
```

## Step 2: Box plot
```{r}
#
```

## Step 3: Histogram
```{r}
#
```

## Step 4: Scatterplot
```{r}
#
```